---
title: "Working with Dates and Times"
author: "Clarissa Boyajian"
date: "8/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Play around with functions in `lubridate`
```{r}
my_date <- "03-15-1998"
class(my_date)
```

Optional orders possible: `ymd()`, `mdy()`, `dmy()`
```{r}
my_date_new <- lubridate::mdy(my_date)
class(my_date_new)
```

#### Try converting the following to dates:

- 4/6/2005 (mdy)
- 20100806 (ymd)
- 12-Jun-1974
- your birthday

```{r}
ex_1 <- "4/6/2005"
print(lubridate::mdy(ex_1))

ex_2 <- "20100806"
print(lubridate::ymd(ex_2))

ex_3 <- "12-Jun-1974"
print(lubridate::dmy(ex_3))

ex_4 <- "May 23, 1990"
print(lubridate::mdy(ex_4))

ex_5 <- c("6/12/1990", "050370")
print(lubridate::mdy(ex_5))
```

#### Times:
Good to use 24 hour clock
```{r}
time <- "2020-08-12 11:18"
ex_5 <- lubridate::ymd_hm(time)
ex_5
#Currently reported in UTC (Uuniversale Greenwich time)
class(ex_5)

time_la <- with_tz(ex_5, "America/Los_Angeles")
time_la
```

#### Practice: 
- create date time
- convert to time in Sydney Australia
```{r}
time_2 <- "2021/08/17 20:46"
time_syd <- with_tz(time_2, "Australia/Sydney")
time_syd
```

#### Pull out week/day/etc. using lubridate
*helpful if you're wanting to add a new column (using `mutate()`) with just a piece of the date - time*
```{r}
lubridate::week(time_syd)
#tells you the week of the year
lubridate::day(time_syd)
#tells you the day of the year
lubridate::minute(time_syd)
```
*find what day of the week a date was one*
```{r}
my_bday <- lubridate::ymd("1990-05-23")
class(my_bday)

wday(my_bday, label = TRUE)
```

#### Manipulating dates in a tibble
```{r}
urchin_counts <- tribble(
  ~date, ~species, ~size_mm,
  "10/3//2020", "purple", 55,
  "10/4/2020", "red", 48,
  "11/17/2020", "red", 67
)

# Have to have info stored as a date class for lubridate functions to work
urchin_counts_ymd <- urchin_counts %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  mutate(year = year(date),
         month = month(date))

```









